\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\markboth{ACKNOWLEDGMENTS}{ACKNOWLEDGMENTS}

This thesis and my whole student years would not have been the same without the many inspiring people that I have met along the way.
Therefore, I would like to express my gratitude to my mentors, colleagues, and friends that have supported me during this journey.

Firstly, I would like to thank my supervisor, Prof.\ Alessandro Ricci for transmitting to me his passion for his research topics through his enthusiasm when teaching, and for giving me the opportunity to live such a unique experience by putting me in contact with some of the brightest people in the field of Multi-Agent Systems.

Then I would like to thank Samuele Burattini for always being there during the development of this thesis, for your reassurance and support, and for your precious advice.
I wish you all the best for a successful and rewarding academic career.

I would also like to thank everyone I had the pleasure to work with in St.\ Gallen while working on this thesis, especially Prof.\ Simon Mayer, Prof.\ Andrei Ciortea, and J\'{e}r\'{e}my Lem\'{e}e for their continuous support and for introducing and guiding me into the world of research.
From the first day I arrived in St.\ Gallen, you made me feel at home and part of your group, and I am truly grateful for that.
Thank you for all the interesting discussions, the lovely lunch chats, and for sharing your knowledge and experience with me.

Speaking of St.\ Gallen, I would also like to thank my flatmate Sebastian and my friends Lukas and Jennifer for making my stay there so enjoyable, letting me have some fun after long days of work, and supporting me when work was overwhelming and through sad and difficult times.
In such a short time, we have been able to share so much and you have become such an important part of my life.

I would wish to thank my family for their unconditional everyday support, the help they always provide me with, and for allowing me to keep studying and following my passions and dreams.
I know that sometimes I am not the most open and outgoing person, but I truly appreciate all the love and care you give me.

Even though probably words are not enough to express my gratitude, I still want to try and thank the members of what I call my second family.
First of all, Simone whom I met during my first year of university and who had the burden to put up with me since.
Hopefully, we can yell at each other some more when working together on other projects in the future.
Equally, I would like to thank Marta and Martina.
We have shared quite possibly all the joy and sorrow of our master's degree.
Thank you for always being by my side and I hope I was only half as good a friend to you as you have been to me.

Last but not least, I would like to thank my colleagues and friends that have been with me during my student years and that have made my time at university so special.


